Accurate binding of calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-09-19 Manuscript/manuscript.tex
Changes Hector

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{csquotes}
\usepackage{mhchem}

@@

According to ECC, electronic
polarizability can be implicitly included in classical MD simulations by
placing all particles into a homogeneous dielectric continuum 
with a dielectric constant $\epsilon _{el}$, 
which is the electronic part of the dielectric constant of 
the media \cite{leontyev11}. Measurements of high frequency 
dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
for almost any biomaterial \cite{some_original_work, leontyev11}.
Such a dielectric continuum can be easily included in standard MD simulation by
a formal transformation of partial charges 
\begin{equation}
  Q^{ECC} = f_q \cdot Q
\end{equation}
with a constant scaling factor $f_q = \epsilon _{el} ^{-1/2}$ 
effectively representing the newly introduced electronic continuum. 
The value measured for water, $\epsilon _{el} = 1.78$, gives 
a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}, which has been
successfully used to improve the performance of force field for ions \cite{kohagen14,kohagen16,??}.

While the scaling factor of $f_q = 0.75$ for ions in water is well justified and 
improves the model performance against experimental scattering data \cite{kohagen14,kohagen16, Pluharova2014},
it is not clear if the same factor should be used for partial charges in molecules,
i.e, lipids in our case. 
Unlike the total charge of molecules, atomic partial charges within each molecule are not experimental observables.
\todo{Text should be clarified from here to the end of the paragraph}
Although there have been proposed several schemes for the assignment of partial charges of molecules, 
methods targetting at fitting of electrostatic potentials are 
most popular among classical fixed charge force fields for biomolecules. \cite{Hu2007}
Currently the most commonly employed scheme is the RESP scheme~\cite{RESP_paper,Singh1984}, 
which fits the electrostatic potential from quantum mechanical calculations. 
The resulting set of fixed partial atomic charges is hence 
the best fit to electrostatic potentials of 
an ensemble of conformations of a molecule. 
Such a physical meaning is especially well suited for the application of ECC, 
which targets at improving electrostatic interactions. 
Depending to the details, partial charge calculations may also partially
include some of the solvent electronic polarizability effects. 
Thus, we expect that
the correct ECC scaling factor, $f_q$, for molecular partial charges
does not have to be striclty $f_q = \epsilon _{el} ^{-1/2}$, 
but rather lies between this value (no electronic polarizability in the charge calculation)
and $1$ (full electronic polarizability already included in the charge calculation).
%
% This is not really important for the paper:
%
%However, the RESP method has several remedies;  %and these are specific to RESP
%RESP charges show a dependence on the orientation and position
%and it does not provide numerically stable partial charges of burried atoms;~\cite{Hu2007}
%the RESP-derived values depend, for example, on the conformational
%sampling of the molecules and description of solvent 
%in the calculations~\cite{RESP_paper, Singh1984, Hu2007, maciejewski14}. 
%
%Hirshfeld-E partitioning: \cite{Verstraelen2013}
%In cases where the partial charges have some physical meaning 
%(e.g. approximate electrostatic potential as in RESP~~\cite{RESP_paper,Singh1984}),
%
%

In this work, we empirically sample this parameter space $f_q \in [0.75, 1.0]$
to find an atomistic MD simulation force field parameters for lipids that
accurately describe the lipid head group response to varying concentrations of cations
when compared to NMR data \cite{akutsu81,altenbach84,scherer89}. Such data can be used to accurately asses
ion binding in PC bilayers, as discussed in Ref, \citenum{catte16} and in section \ref{section:electrometer}.
We chose Lipid14~\cite{dickson14} force field parameters as a starting point,
because their response to bound ions was apparently most realistic against NMR data
in recent work by NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
Also glycerol backbone and head group structures in Lipid14 model were
relatively realistic when compared with other state of the art lipid models \cite{botan15}.
The ECC correction was applied to Lipid14 parameters by scaling partial charges of the head group,
glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
are expected to have the largest contribution to the cation binding.
The hydrocarbon chain parameters are not modified, because they are
already highly optimized and give generally a good description for
hydrophobic part of lipid bilayers in various conditions
in most lipids models, including Lipid14 \cite{ollila16}.
In contrast, improvements in glycerol backbone and head group
parameters are required in all available lipid models~\cite{botan15}.

Exploring different scaling factor values, applied to head group, glycerol backbone
and carbonyl charges of Lipid14 model, we found out that ion binding and
related head group order parameter responses become weaker. 
The optimal behaviour of ion binding was observed with
the scaling factor $f_q = 0.8$. % is optimal for applying ECC on Lipid14 model. 
Interestingly, this scaling factor is in line with the estimate
given by ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013}
combined with RESP calculations in vacuum and implicit solvent reported
in \cite{maciejewski14}. IPolQ charges are obtained as the average of
partial charges given by RESP calculation~\cite{RESP_paper}
in vacuum and in a solvent. Applying the scaling factor of 0.75 to
IPolQ charges calculated from the data in Ref.~\cite{maciejewski14},
gives similar partial charges to ones obtained by scaling Lipid14 charges
with a factor 0.8.
%-----\footnote----------
  %{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
  %types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
  %partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
  %I use 0.80 as an approximate round value, which accounts for the already lower charges in Lipid14. 
  %The use of 0.78---.79 might be more appropriate, though.}

While, the charge scaling improved the behaviour of lipid-ion interactions,
it reduced the area per molecule of lipid bilayer without ions below experimental
values. Simulations with Lipid14 parameters having partial charges of head group, glycerol
backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$,
which is significantly smaller than the experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
and the original Lipid14 value $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
The decrease of area was found to arise from a lower hydration of the lipid head group region,
which can be explained by the increased solvation free energy due to the lower polarity
of molecules with scaled charges. The hydration can be increased
by reducing the effective radius of atoms by changing the 
the $\sigma$ parameters in Lennard-Jones potential for the selected atoms
similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
This decreases the solvation free energy by allowing water molecules to
approach closer to lipid atoms and have stonger electrostatic interactions with them.
After reducing the $\sigma$ parameters with a factor of $f_\sigma = 0.89$
for the same atoms for which charges were scaled, the area per molecule value was
again in agreement with experimental value (see Table~\ref{tab:apls}).

@@

\subsection{Ion concentrations and binding affinity}
The salt concentrations are reported in two different ways
in the used experimental data for the lipid head group order
parameters. The data for DPPC is reported
by using solution concentrations before solvating
the lipids~\cite{akutsu81}, while the data
for POPC~\cite{altenbach84} is reported by using
supernatant concetration measured with atomic absorption
spectroscopy after solvation of lipids.
The former was used in comparison with experiments and
calculated from simulations in previous work as 
[salt]=N$_{\rm c} \times$[water]\,/\,N$_{\rm w}$, where [water]\,=\,55.5~M~\cite{catte16}.
Here we use the latter, because the focus is on POPC bilayers.
This is determined from the ion concentrations, which reach
a constant value at the edges of large enough simulation box.
The two concentrations differ somewhat for CaCl$_2$ systems, because some
amount of ions bind in lipids. However, the small difference do
not affect the conclusions in this or in previous work~\cite{catte16}.

The ion concentrations with constant value at the simulation box edges
are also used to quantify the ion binding affinities is different simulation
models by using the relative surface excess of ions with respect to water $\Gamma_i^w$.
This is defined as \cite{chattorajBOOK}
\begin{eqnarray}
  \Gamma_i^w   &=& \Gamma_i-\Gamma_w \frac{C_i-C_i'}{C_w-C_w'} \\
  \Gamma_{a} &=& \frac{n_{a} - C_{a} V - C_{a}' V'}{A} \, \, ; \, a=i,w \, ,
\end{eqnarray}
where index $i$ and $w$ denote ions and water respectively,
prime accent denotes the phase inside bilayer, 
whereas non-prime symbols account for bulk water phase,
$n$ is the total amount of water or ions,
$C$ denotes bulk concentrations, 
$V$ is the volume of the respective phase separated by the Gibbs divinding plane, 
$A$ is the area of the interface, 
and $\Gamma_a$ is the surface excess of water or ions. 
As $\Gamma_i^w$ does not depend on the definition of the interfacial region
and the position of the Gibbs dividing plane~\cite{chattorajBOOK},
we consider the whole simulation box as an interface. 
Because bilayers have an interface
in both leaflefts,
the total area of the interface in the simulation
 is given by the simulation box area multiplied by two, $2A_b$.
The concentrations of water and ions inside
the bilayer can be expected to vanish, thus $c_{i,w}'=0$.
This provides us a simplified relation for 
the relative surface excess of lipid bilayers in simulations
\begin{equation}\label{surfexcess}
  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
\end{equation}
where the bulk concentrations, $C_w$ and $C_i$, are 
concentrations with constant value at the simulation box edges.

@@

Lipid bilayer structure without ions was validated against NMR experiments
by order parameters for C-H bonds and \mbox{x-ray} scattering experiments by using
form factors. The former validates the structures sampled by individual
lipid molecules in simulations with segmental resolution, while the latter
validates the dimensions of the lipid bilayer (thickness and area per
molecule)~\cite{ollila16}.

The order parameters were calculated from simulations for all C-H bonds
in lipid molecules by using Eq. \ref{OP}. Form factors were calculated 
from equation
\begin{equation}
  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
\end{equation}
where $f_\alpha(q_z)$ is the density of atomic scattering length, 
$\rho_s$ is the density of solvent scattering length,
$n_\alpha (z)$ is the number density of atom $\alpha$ and
$z$ is the distance from the membrane centre along its normal 
spanning the membrane with thickness $D$.

@@

\subsubsection{Simulations with aqueous ions}
The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
(POPC) bilayer and an aqueous salt solutions of varying concentrations.
In particular, the periodic orthorhombic simulation box conatined 
128 POPC molecules and approximately 50~water molecules per each lipid.
Water molecules were described by OPC3 model~\cite{Izadi16} 
as it is currently the most accurate three site rigid water model. 
In order to test transferability of the newly developed ECC-lipids model, 
we also performed several additional simulations with water models 
OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005}
presented in Supporting Information (SI). 
We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
The simulation settings used in this work are 
summarized in Table \ref{tbl:mdpar}, 
and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
  this is done in Amber by default. If this is correct, the information in the
  table do not match with this sentence. Based on semi-extensive testing I made few years
ago this do not change anything. Anyway, this should be reported consistently.}
Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.

@@

\subsubsection{Simulations with cationic surfactants}
Automated topology builder \cite{malde11} was first used to create the structure of
dihexadecyldimethylammonium bromide, C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$, molecule.
AmberTools program \cite{amber} was then used to generate the Amber-type force field
parameters. The parameters were converted to the Gromacs format by using
acpype tool \cite{acpype}. The partial charges were then manually modified
to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
The surfactants were randomly placed among the lipids to form bilayer structures with
mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
6, 14, 21, 35 or 50 surfactants per leaflet.
The systems were simulated for 200~ns using Lipid14 model for POPC.
First 20~ns were omitted from the analysis.

@@

\subsection{POPC membrane and its structure}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \caption{\label{simVSexpNOions}
    Top: chemical structure of POPC with definitions of headgroup order parameters.
    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
    compared with experimental order parameters from \cite{ferreira13}.
    Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
    }
\end{figure}

@@

In order to validate the newly developed model, ECC-lipids, 
we compared our simulation results without any ions to
NMR order parameters measurements and x-ray scattering form factors 
(Fig.~\ref{simVSexpNOions} and Table~\ref{tab:apls}). 
The tail order parameters being already highly optimized in the original Lipid14 model~\cite{dickson14}
are found to match the experimental values nicely.
\todo{The discussion about acyl chain and headgroup order parameters needs to be
  finished when the value for C2, the scale are fixed in the figure and
  size of the points are fixed in the figure.}
The headgroup order parameters are slightly larger in the ECC-lipid model,
which is in line with larger P-N vector angle in Fig.~\ref{OrderParameterCHANGESsurf}.
With the current data we cannot, however, conclude which is the more realistic
conformation since in both models other headgroup order parameter agrees
with experiments within error bars and other does not.
Slight changes are also observed in glycerol backbone order parameters,
bu the accuracy of both models is comparable to the state of art lipid models
available in literature \cite{ollila16}. The agreement between the x-ray scattering
form factors and the areas per molecule from simulations and experiments
confirm that the membrane structural properties are well captured. 
A structural comparison of ECC-lipids with Lipid14 can be found 
in SI along with results with other water models. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

@@

The headgroup order parameter changes with increasing amount of 
cationic surfactant dihexadecyldimethylammonium bromide (C$_{12}$Cl$_{16}$$^+$N2C$_1$Br$^-$) 
is compared between experiments~\cite{scherer89} and simulations in
Fig. \ref{OrderParameterCHANGESsurf}. 
The observed order parameter decrease is simulations and experiments
can be approximated to be linear at least with the mole fractions
below $\sim$30\%, as expected from Eq. \ref{OPchangeEQ}.
The slope is, however, too steep in Lipid14 model indicating that 
the head group order parameters are too sensitive to a bound charge.
The ECC-lipids model gives a slope in very good agreement with experiments
for the $\alpha$ segment, while the slope is slightly
underestimated for the $\beta$ segment.
%This has to be taken into account when interpreting the
%head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
%observed in  Ref. \citenum{catte16}, see also Fig. \ref{OrderParameterCHANGESnewMODELS}.
%The present comparison of head group order parameter changes 
%suggests that the observed overestimated response of all models to 
%increasing CaCl$_2$ concentration in \cite{catte16} 
%can be explained at least in part by a high sensitivity of the head group order parameter response
%to the bound charge.
\todo{SAMULI: We could calculate the slopes from simulations, but I am not sure if
we would actually learn anything useful from this.}

The headgroup P-N vector angle with respect to the membrane normal
is also shown as a function of cationic surfactant mole fraction
in Fig.~\ref{OrderParameterCHANGESsurf}.
The headgroup orients more parallel to the membrane
normal with increasing amount of bound cations, as expected~\cite{seelig87}. 
The effect is more pronounced in Lipid14 than 
in ECC-lipids model, which is in line with the order parameter 
results and the reduced charge--dipole interactions in the ECC-lipid model.
The response of $\alpha$-order parameter to bound positive charge
in ECC-lipid model is in good agreement with experiments.
The model can be thus used to study changes of lipid P-N vector
in varying conditions.

>>>>

\usepackage{rotating} 
\usepackage{times}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage{amsmath}
\usepackage{epstopdf}
\usepackage[obeyFinal]{easy-todo}
\usepackage{csquotes}
\usepackage{mhchem}
\usepackage{chemfig}

@@

According to ECC, electronic
polarizability can be implicitly included in classical MD simulations by
placing all particles into a homogeneous dielectric continuum 
with a dielectric constant $\epsilon _{el}$, 
which is the electronic part of the dielectric constant of 
the media \cite{leontyev11}. Measurements of high frequency 
dielectric constant gives values of approximately $\epsilon _{el} \approx 2$ 
for almost any biomaterial \cite{some_original_work, leontyev11}.
Such a dielectric continuum can be easily included in standard MD simulation by
a formal transformation of partial charges 
\begin{equation}
  Q^{ECC} = f_q \cdot Q
\end{equation}
with a constant scaling factor $f_q = \epsilon _{el} ^{-1/2}$ 
effectively representing the newly introduced electronic continuum. 
Assuming globally a  high frequency dielectric constant as measured in water (square of the refraction index),  $\epsilon _{el} = 1.78$, results 
in a scaling factor of $f_q = 0.75$ \cite{some_orig_source, leontyev11}. This scaling factor has been
successfully used to improve the performance of force field for ions in solution \cite{kohagen14,kohagen16,??} which then agree with neutron scattering data \cite{kohagen14,kohagen16, Pluharova2014}.

While the scaling factor of $f_q = 0.75$ for ions in water improves their performance and is physically justifiable within the MDECC theory \cite{Leontyev2011},
it is not clear how the same factor should be used for partial charges in molecules,
e.g., lipids in our case. Unlike the total charge of molecules, atomic partial charges within each molecule are not physical observables.
There are several schemes for the assignment of partial charges for biomolecules. \cite{Hu2007}
Currently, the most commonly employed scheme is RESP~\cite{RESP_paper,Singh1984}, 
which aims to reproduce the electrostatic potential obtained from quantum mechanical calculations around an ensemble the target molecules. 
The fixed partial atomic charges obtained by this method is hence 
the best fit of these partial charges that reproduces the electrostatic potential around the provided target conformations. 
Such averaging is well suited for the application of ECC, 
which targets at improving electrostatic interactions using a mean field correction. 
In practice, partial charges currently implemented in force fields may already 
include  to some extend some of the solvent electronic polarizability effects, i.e.,
the RESP charges are scaled to fit some experimental observable. 
Thus, we expect that the application of the ECC scaling factor, $f_q$, on the molecular partial charges
included in the available force fields do not necessarily require the integral, $f_q = \epsilon _{el} ^{-1/2}$, 
but instead it lies between 0.75 (no electronic polarizability in the charge calculation)
and $1$ (full electronic polarizability already included in the charge calculation).
%
% This is not really important for the paper:
%
%However, the RESP method has several remedies;  %and these are specific to RESP
%RESP charges show a dependence on the orientation and position
%and it does not provide numerically stable partial charges of burried atoms;~\cite{Hu2007}
%the RESP-derived values depend, for example, on the conformational
%sampling of the molecules and description of solvent 
%in the calculations~\cite{RESP_paper, Singh1984, Hu2007, maciejewski14}. 
%
%Hirshfeld-E partitioning: \cite{Verstraelen2013}
%In cases where the partial charges have some physical meaning 
%(e.g. approximate electrostatic potential as in RESP~~\cite{RESP_paper,Singh1984}),
%
%

In this work, we produce a phospholipid model for classical MD simulations that
accurately describes the lipid head group response to varying concentrations
of monovalent and divalent cations. This is critical membrane feature, which is
poorly reproduced by currently available models, can affect not only membrane properties
in the presence of ions but also modulates the interaction with charged moieties in the surface.
Importantly, this response from simulations can be accurately
compared against experimental NMR data \cite{akutsu81,altenbach84,scherer89},
as discussed in Ref. \citenum{catte16} in section \ref{section:electrometer}.
To this end, we empirically explore the scaling factor parameter space, $f_q \in [0.75, 1.0]$ on the Lipid14~\cite{dickson14} force field.
We selected this force field as a starting point because their response to bound ions was apparently the most realistic against NMR data
in recent work by NMRlipids project (see Fig.~5 in Ref.~\citenum{catte16}).
Also glycerol backbone and head group structures in Lipid14 model were
relatively realistic when compared with other state of the art lipid models \cite{botan15}.
The ECC correction was applied to Lipid14 parameters by scaling partial charges of the head group,
glycerol backbone and carbonyl regions, which are the most polar parts in lipids and
are expected to have the largest contribution to the cation binding \todo{Do we really need to scale charges in non charged methylens in the choline and glycerol backbones? why? Does it really make a difference?}.
However, we do not modify the hydrocarbon chain parameters, because they are
already highly optimized and give generally a good description for
hydrophobic part of lipid bilayers in most lipids models, including Lipid14 \cite{ollila16}.
This contrast with the behavior in glycerol backbone and head group
parameters which call for improvements in all available lipid models~\cite{botan15}.

Exploring different scaling factor values, we found out that ion binding and
related head group order parameter responses become weaker. 
The optimal behaviour of ion binding was observed with
the scaling factor $f_q = 0.8$. % is optimal for applying ECC on Lipid14 model. 
Interestingly, this scaling factor is in line with the estimate
given by ``implicitly polarized charges'' (IPolQ) \cite{ipolq2013}
combined with RESP calculations in vacuum and implicit solvent reported
in \cite{maciejewski14}. IPolQ charges are obtained as the average of
partial charges given by RESP calculation~\cite{RESP_paper}
in vacuum and in a solvent. Applying the scaling factor of 0.75 to
IPolQ charges calculated from the data in Ref.~\cite{maciejewski14},
gives similar partial charges to ones obtained by scaling Lipid14 charges
with a factor 0.8.
%-----\footnote----------
  %{Depending on which QM method you arrive at values from 0.76 to 0.83, averages across atom
  %types being around 0.78--0.80. Even though the methods are almost identical, authors of Lipid 14 find lower
  %partial charges in vacuum than here -- so I prefer the higher value. As the choice of charges is arbitrary anyway,
  %I use 0.80 as an approximate round value, which accounts for the already lower charges in Lipid14. 
  %The use of 0.78---.79 might be more appropriate, though.}

While, the charge scaling improved the behaviour of lipid-ion interactions,
it reduced the area per molecule of lipid bilayer without ions below experimental
values. Simulations with Lipid14 parameters having partial charges of head group, glycerol
backbone and carbonyls scaled with 0.8 gave the area per molecule value of $\approx$60~\AA$^2$,
which is significantly smaller than the experimental value 64.3~\AA$^2$ (\cite{}\todo{missing REF for APL experiment})
and the original Lipid14 value $(65.6 \pm 0.5)$~\AA$^2$~\cite{dickson14}.
The decrease of area was found to arise from a lower hydration of the lipid head group region,
which can be explained by the increased solvation free energy due to the lower polarity
of molecules with scaled charges. The hydration can be increased
by reducing the effective radius of atoms by changing the 
the $\sigma$ parameters in Lennard-Jones potential for the selected atoms
similarly as done for free ions in solution \cite{kohagen14,kohagen16,Pluharova2014}.
\todo{We should discuss how this can potentially affect the intermolecular interaction when mixing scaled and non scaled molecules}
This decreases the solvation free energy by allowing water molecules to
approach closer to lipid atoms and have stonger electrostatic interactions with them.
After reducing the $\sigma$ parameters with a factor of $f_\sigma = 0.89$
for the same atoms for which charges were scaled, the area per molecule value was
again in agreement with experimental value (see Table~\ref{tab:apls}).

@@

\subsection{Salt concentrations and binding affinity}
Experimental works providing the lipid head group order
parameters report the salt concentrations in two different ways.
For DPPC, salt concentrations before solvating
the lipids are used~\cite{akutsu81}. Instead, for POPC,
the salt concentration of supernatant after solvation is
measured using atomic absorption~\cite{altenbach84}.
In this work, we will use the latter definition. For this we
will measure the salt concentration in the water bulk region,
\textit{i.e.}, the farthest point from both lipid leaflets in the water phase. 
Note that in our previous study, Ref.~\citenum{catte16},
we calculated the concentration as in the DPPC experiments.
Despite the two concentrations may differ significantly for CaCl$_2$ systems
these differences do not affect qualitatively the conclusions neither in this
nor in our previous work~\cite{catte16}.

To quantify the ion binding affinity to a membrane, we
calculated the relative surface excess of ions with respect to water $\Gamma_i^w$  \cite{chattorajBOOK}.
This quantity does not depend on the definition of the interfacial region
and the position of the Gibbs dividing plane. Therefore, we consider our whole simulation box as
an interface flanked by a water phase and an ideal phase with the characteristics of to
the hydrophobic region in the middle of the membrane where no ions or waters are present.
This setup provides a simplified relation for $\Gamma_i^w$ in lipid bilayers simulations 
\begin{equation}\label{surfexcess}
  \Gamma_i^w=\frac{1}{2A_b} \left ( n_i - n_w \frac{C_i}{C_w} \right ) ,
\end{equation}
where $n_w$ and $n_i$ are the total number of waters and ions in the system;
$C_w$ and $C_i$ are their respective bulk concentrations in the water phase;
and $A_b$ is the size of the box in the membrane plane. Therefore, this quantity
does not depend on any arbitrary definition of a bound ion to a membrane and
therefore allow us to accurately compare between different models. Note that
because bilayers have an interface in both leaflefts, the total area of the
interface is twice the area of the membrane, $2A_b$.

@@

Lipid bilayer structure without ions was validated against NMR and 
\mbox{x-ray} scattering experiments by calculating order parameters
for C-H bonds and form factors from our simulations. The former validates the structures sampled by individual
lipid molecules in simulations with atomic resolution, while the latter
validates the dimensions of the lipid bilayer (thickness and area per
molecule)~\cite{ollila16}.

The order parameters were calculated from simulations for all C-H bonds
in lipid molecules by using Eq. \ref{OP}. Form factors were calculated 
from equation
\begin{equation}
  F(q) = \int _{-D/2} ^{D/2} \left ( \sum _\alpha f_\alpha (q_z) n_\alpha (z) - \rho _s \right ) \exp (izq_z) \mathrm{d}z,
\end{equation}
where $f_\alpha(q_z)$ is the density of atomic scattering length, 
$\rho_s$ is the density of solvent scattering length in the bulk region,
$n_\alpha (z)$ is the number density of atom $\alpha$ and
$z$ is the distance from the membrane centre along its normal 
spanning until the water bulk region, $D$.

@@

\subsubsection{Simulations of POPC bilayers in aqueous ions}
The simulated systems consisted of 1-Palmitoyl-2-oleoylphosphatidylcholine
(POPC) bilayer and in aqueous solutions of varying salt concentrations.
In particular, the periodic orthorhombic simulation box contained 
128 POPC molecules and approximately 50~water molecules per each lipid.
Our default, water molecules were described by OPC3 model~\cite{Izadi16} 
which is currently the most accurate three site rigid water model. 
In order to test transferability of our newly developed ECC-lipids model, 
we also performed several additional simulations with water models 
OPC \cite{Izadi14}, SPC/E \cite{Berendsen1987}, TIP3p-FB and TIP4p-FB~\cite{Wang2014}, and TIP4p/2005 \cite{Abascal2005}
presented in Supporting Information (SI). 
We used ECC-ions model for ions \cite{jungwirth17-new-paper-to-be-published, kohagen16, Pluharova2014}.  % cite order: [Ca_2s, Na_s, Cl_2s]
Simulations with Lipid14 use ion models by Dang~\cite{smith94,chang1999,dang2006}, and by \AA{}qvist~\cite{aqvist90}. 
Classical molecular dynamics simulations were performed using the GROMACS \cite{Abraham15} simulation package (version 5.1.4).  
The simulation settings used in this work are 
summarized in Table \ref{tbl:mdpar}, 
and they are based on previously used settings in \cite{catte16} available at \cite{lipid14POPC0mMNaClfiles}.
\todo{As far as remember, I used there Langevin dynamics instead of thermostated MD, because
  this is done in Amber by default. If this is correct, the information in the
  table do not match with this sentence. Based on semi-extensive testing I made few years
ago this do not change anything. Anyway, this should be reported consistently.}
Simulation trajectories and parameters are available at \cite{??} \todo{To be uploaded to Zenodo}.

@@

\subsubsection{Simulations of POPC bilayers with cationic surfactants}
Automated topology builder \cite{malde11} was first used to create the structure of
dihexadecyldimethylammonium bromide cationic surfactant,

\vspace{0.5cm}
\chemfig{
\chemabove{N}{\scriptstyle\oplus} (-[:135]H_3C)(-[:225]H_3C)(-[:315]{(}CH2{)}_{16}-CH_3)(-[:45]{(}CH2{)}_{12}-CH_3)  (-[:0,3.5,,,draw=none]\chemabove{Br}{\scriptstyle\ominus}) 
} .
\vspace{0.5cm}

AmberTools program \cite{amber} was then used to generate the Amber-type force field
parameters. The parameters were converted to the Gromacs format by using
acpype tool \cite{acpype}. The partial charges were then manually modified
to approximately correspond to their equivalent segments in Lipid14 \cite{dickson14}.
The surfactants were randomly placed among the lipids to form bilayer structures with
mole fractions 10\%, 20\%, 30\%, 42\% and 50\% of surfactant in the POPC bilayer.
All systems contained 50 POPC molecules per leaflet, 6340 TIP3P water molecules and
6, 14, 21, 35 or 50 surfactants per leaflet.
The systems were simulated for 200~ns using Lipid14 model for POPC where reasonable lipid neighbor exchange occurs.
First 20~ns were omitted from the analysis.

@@

\subsection{POPC membrane structure and dynamics}

\begin{figure}[tbp]
  \centering
  \includegraphics[width=7.5cm]{../Fig/POPCstructure.eps}
  \includegraphics[width=8.0cm]{../Fig/ipython_nb/Order-parameters_exp-L14-ECCL17_q80_sig89.eps}
  \includegraphics[height=8.4cm,angle=-90]{../Fig/form-f_exp-l14-eccl17.eps}
  \caption{\label{simVSexpNOions}
    Top: Chemical structure of POPC with definitions of different order parameters calculated.
    Middle: Order parameters of head group, glycerol backbone and sn-1 and sn-2 tails  from simulations with Lipid14 \cite{dickson14} and ECC-lipids models
    compared with experimental order parameters from \cite{ferreira13} {\color {red} X-label misses glycerol region}.
    Bottom: X-ray scattering form factors from experiments~\cite{Kucerka2011} and simulations using Lipid14 \cite{dickson14} and ECC-lipids models. 
    }
\end{figure}

@@

In order to validate the newly developed model, ECC-lipids, 
we compared our simulation results without any ions to
NMR order parameters measurements and x-ray scattering form factors 
(Fig.~\ref{simVSexpNOions}). 
Our tail order parameters, as in the original Lipid14 model~\cite{dickson14},
match the experimental values. This is expected as the ECC model does not modify
the already highly optimized parameters of this region in Lipid14. 
\todo{The discussion about acyl chain and headgroup order parameters needs to be
  finished when the value for C2, the scale are fixed in the figure and
  size of the points are fixed in the figure.}
The headgroup order parameters computed form our new are slightly larger in the ECC-lipid model in comparison ,
which is in line with larger P-N vector angle in Fig.~\ref{OrderParameterCHANGESsurf}.
With the current data we cannot, however, conclude which is the more realistic
conformation since in both models other headgroup order parameter agrees
with experiments within error bars and other does not.
Slight changes are also observed in glycerol backbone order parameters,
bu the accuracy of both models is comparable to the state of art lipid models
available in literature \cite{ollila16}. The agreement between the x-ray scattering
form factors (Fig.~\ref{simVSexpNOions}) and the areas per molecule from simulations and experiments (Table~\ref{tab:apls}) 
confirm that the membrane structural properties are well captured. 
A structural comparison of ECC-lipids with Lipid14 can be found 
in SI along with results with other water models. 
%The optimal value for $f_\sigma$ was found by representing the overall membrane structure well
%by matching scattering form factors to experimental data~\cite{Petrache06, Kucerka08, Pabst10}.

\todo{Dynamics check is missing: MSD (Hector/Joe)}

@@

The headgroup order parameter changes with increasing amount of 
cationic surfactant dihexadecyldimethylammonium bromide 
is compared between experiments~\cite{scherer89} and simulations in
Fig. \ref{OrderParameterCHANGESsurf}. 
The observed order parameter decrease is simulations and experiments
can be approximated to be linear at least with the mole fractions
below $\sim$30\%, as expected from Eq. \ref{OPchangeEQ}.
The slope is, however, too steep in Lipid14 model indicating that 
the head group order parameters are too sensitive to a bound charge.
The ECC-lipids model gives a slope in very good agreement with experiments
for the $\alpha$ segment, while the slope is slightly
underestimated for the $\beta$ segment.
%This has to be taken into account when interpreting the
%head group order parameter response to CaCl$_2$ concentration in Lipid14 model.
%observed in  Ref. \citenum{catte16}, see also Fig. \ref{OrderParameterCHANGESnewMODELS}.
%The present comparison of head group order parameter changes 
%suggests that the observed overestimated response of all models to 
%increasing CaCl$_2$ concentration in \cite{catte16} 
%can be explained at least in part by a high sensitivity of the head group order parameter response
%to the bound charge.
\todo{SAMULI: We could calculate the slopes from simulations, but I am not sure if
we would actually learn anything useful from this.}

The headgroup P-N vector angle with respect to the membrane normal
is also shown as a function of cationic surfactant mole fraction
in Fig.~\ref{OrderParameterCHANGESsurf}.
The headgroup orients more towards the water phase with increasing
amount of bound cations, as previously reported in Ref.~\citenum{seelig87}. 
The effect is more pronounced in Lipid14 than 
in ECC-lipids model, which is in line with the order parameter 
results and the reduced charge--dipole interactions in the ECC-lipid model.
The response of $\alpha$-order parameter to bound positive charge
in ECC-lipid model is in good agreement with experiments.
The model can be thus used to study changes of lipid P-N vector
in varying conditions.
